---
title: "Homework 2"
author: Alexis Colangelo
date: 10/3/2019
output: github_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1: Mr. Trash Wheel

### Read and clean Mr. Trash Wheel sheet
 * Specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel.`
 * Use reasonable variable names
 * Omit rows that do not include dumpster-specific data
 * Round the number of sports balls to the nearest integer and  convert the result to an integer variable (using as.integer)

```{r}
trash_data =
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 1, range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(sports_balls = as.integer(sports_balls))
  
```

### Read and clean precipitation data for 2017 and 2018. 
 * For each, omit rows without precipitation data and add a variable year. 

```{r}
precip_17 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
```

```{r}
precip_18 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)
```

* Combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
precip_data =
  full_join(precip_17, precip_18, by = c("month", "year", "total")) %>%
  arrange(month) %>%
  mutate(month = month.name[month])
```

### Write a paragraph about these data; you are encouraged to use inline R. 
* Be sure to note the number of observations in both resulting datasets, and give examples of key variables.
* For available data, what was the total precipitation in 2018? 
* What was the median number of sports balls in a dumpster in 2017?

In the Mr. Trash Wheel data set, there are `r nrow(trash_data)` observations and `r ncol(trash_data)` columns or variables for each observation. An example of a key variable includes the weight of trash removed, which totals `r sum(pull(trash_data, weight_tons))` tons from May 2014 to June 2019. Another key variable is homes powered, which the average dumpster powered `r mean(pull(trash_data, homes_powered))` homes in the same time frame. The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(trash_data, year == 2017), sports_balls))`. 

In the precipitation data set, there are `r nrow(precip_data)` observations and `r ncol(precip_data)` columns or variables for each observation. An example of a key variable is total precipitation, which averaged `r mean(pull(filter(precip_data, year == 2017), total))` in 2017. An example of another key variable is month, the average precipitation in July for 2017 and 2018 was `r mean(pull(filter(precip_data, month == "July"), total))`. The total precipitation in 2018 was `r sum(pull(filter(precip_data, year == 2018),total))`. 

## Problem 2: 538 

### Clean the data in pols-month.csv.

* Use separate() to break up the variable mon into integer variables year, month, and day
* Replace month number with month name
* Create a president variable taking values gop and dem, and remove prez_dem and prez_gop;
* remove the day variable.

```{r, message=FALSE}
pols_data =
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = month.name[month]) %>%
  mutate(prez_gop = as.character(prez_gop)) %>%
  mutate(prez_gop = recode(prez_gop, "1" = "gop", "0" = "dem")) %>%
  mutate(prez_dem = as.character(prez_dem)) %>%
  mutate(prez_dem = recode(prez_dem, "1" = "dem", "0" = "gop")) %>%
  mutate(prez = prez_gop) %>%
  select(-prez_gop, -prez_dem, -day)
```

### Clean the data in snp.csv

* Clean the data in snp.csv using a similar process to the above
* For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r, message=FALSE}
snp_data =
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = month.name[month]) %>%
  arrange(year, month) %>%
  select(year, month, close, -day)
```

### Tidy the unemployment data so that it can be merged with the previous datasets. 
* This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r, message=FALSE}
unemploy_data =
   read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemploy_percent") %>%
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))
```

### Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r, message=FALSE}
fte_data =
  left_join(pols_data, snp_data, by = c("month","year")) %>%
  mutate(year = as.numeric(year)) %>%
  left_join(unemploy_data, by = c("month", "year"))
```

### Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

The first dataset (pols_data) contained information from January 1947 to June 2015 on which political party held the presidency, and the number and political party of senators, governors, and representatives at that particular time. Pols_data contained `r nrow(pols_data)` total observations and `r ncol(pols_data)` columns/variables. The second dataset (snp_data) included S&P stock index values at close from April 1950 to May 2015. Snp_data contained `r nrow(snp_data)` total observations and `r ncol(snp_data)` columns/variables. Months and years were included in the dataset. The third dataset (unemploy_data) contained unemployment rates in percents for each month and year observed ranging from January 1948 to June  2015 (July to December 2015 rates were NA in the dataset). Unemploy_data contained `r nrow(unemploy_data)` total observations and `r ncol(unemploy_data)` columns/variables. 

The resulting dataset (fte_data) containing the three datasets mentioned above contained information ranging from January 1947 to June 2015. It included `r nrow(fte_data)` total observations and `r ncol(fte_data)` columns/variables. Some key variables in the total dataset include the political party of the sitting president, stock index value at close, the percent of unemployment, months, and years.

## Problem 3: Baby Names

### Load and tidy the data. 
* Although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. 

* Also, some rows seem duplicated, and these will need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).

```{r, message=FALSE}
name_data =
  read_csv(file = "./data/popular_baby_names.csv") 
```

%>%