---
title: "Homework 2"
author: Alexis Colangelo
date: 10/3/2019
output: github_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1: Mr. Trash Wheel

### Read and clean Mr. Trash Wheel sheet
 * Specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel.`
 * Use reasonable variable names
 * Omit rows that do not include dumpster-specific data
 * Round the number of sports balls to the nearest integer and  convert the result to an integer variable (using as.integer)

```{r}
trash_data =
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 1, range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(sports_balls = as.integer(sports_balls))
  
```

### Read and clean precipitation data for 2017 and 2018. 
 * For each, omit rows without precipitation data and add a variable year. 

```{r}
precip_17 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
```

```{r}
precip_18 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)
```

* Combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
precip_data =
  full_join(precip_17, precip_18, by = c("month", "year", "total")) %>%
  arrange(month) %>%
  mutate(month = month.name[month])
```



